---
title: "Suicide Rates"
author: "Michelle Wong, Ploypilin Pruekcharoen, Yudong Li, Adam Ahmed"
date: "2/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("summary.R")
source("summary_table.R")
source("facility_chart.R")
source("suicide_age_chart.R")
source("suicide_gdp_chart.R")
```

# Suicide Rates in Countries

    An introduction of the problem domain and dataset (feel free to borrow language you used in your brainstorm assignment)
    A paragraph of summary information, citing at least 5 values calculated from the data
    A table of  aggregated data (e.g., that has been "grouped" by one of your features)
    Three charts that display information from the data (what you visualize is up to you). 
index.Rmd File

In the Markdown document you create, you should leverage the functions written in the other files to create your report:

    Write an introductory paragraph explaining the domain/data set (feel free to borrow from the brainstorm part of the assignment)
    Write a summary paragraph that includes information calculated by your summary information function
    Render your table and describe pertinent observations found in it
    Describe the purpose of each chart (e.g., "this chart was intended to show the relationship between..."), render each chart, and describe pertinent observations found in each chart

Each function must return a different type of visualization (e.g., you can only create one Scatter Plot).  Again, in your report you must describe why you included the chart (e.g., what it attempts to seeks to express), and what information it reveals. For each chart, we expect the following:

    Proper labels/titles/legends
    Intentional chart type and encoding selection based on the question of interest and data type

# Suicide Rates

## Introduction
Our project will be investigating suicide rates globally in comparison to the mental health facilities offered in each country, and the GDP per Capita of each country. All data is taken for the most recent years that the country had data for the category. Suicide has become more widely talked about in social media in recent years, however it seems that the awareness of mental health has been an emerging topic in developed countries but is not as emphasized in undeveloped countries. Suicide has resulted in the loss of many amazing people. As such, we want to investigate possible correlations between these different indicators, hopefully determining a reason or a trend for higher suicide rates in particular countries over others.

We will be using combined data from both [Suicide Rates Overview 1985 to 2016](https://www.kaggle.com/russellyates88/suicide-rates-overview-1985-to-2016) and [Mental health service availability](http://apps.who.int/gho/data/node.main.MHFAC?lang=en) to create data visualizations that may aid us in finding these trends. 

Our combined dataset has `r countries_no` countries and columns: `r col_names`. The average suicide rate (per 100k population) was `r  average_suicide_rate`, with the largest suicide rate of `r max_suicide_rate`. The average GDP per capita for this dataset is `r average_gdp` and the average number of mental health facilities offered (per 100k population) was `r average_mental_hospitals`.

## Summary Table

This table is sorted by... and contains all countries we are assessing as well as their corresponding average GDPs per capita, number of mental health facilities, and the total number of suicides across the years.

``` {r, echo=FALSE}
get_summary_information()
```

From this summary, we can observe that...

---

## Mental Health Facilities Available

This map displays the mental health facilities available globally, showing the 
```{r, echo=FALSE}
source("facility_chart.r")
world_shape <- create_map()
 blank_theme <- theme_bw() +
   theme(
     axis.line = element_blank(),        # remove axis lines
     axis.text = element_blank(),        # remove axis labels
     axis.ticks = element_blank(),       # remove axis ticks
     axis.title = element_blank(),       # remove axis titles
     plot.background = element_blank(),  # remove gray background
     panel.grid.major = element_blank(), # remove major grid lines
     panel.grid.minor = element_blank(), # remove minor grid lines
     panel.border = element_blank()      # remove border around plot
   )
 
 ggplot(world_shape) + 
   geom_polygon(
     mapping = aes(x = long, y = lat, group = group, fill = total_avalibility),
     color = "white", 
     size = .1        
   ) + 
   coord_map(xlim=c(-180,180)) + 
   scale_fill_continuous(low = "#cc3300", high = "#800000") + 
   labs(fill = "total_avalibility") + 
   blank_theme

create_map
```

## Suicide Rates, Gender, and Age

This bar chart displays the correlation between the different age groups of each gender (Male or Female) for each country, displaying the trend.

``` {r, echo=FALSE}
chart1
```
From this chart, we can observe a ...

## Suicide Rates, GDP

We created a scatterplot to display the GDP per capita of each country against their respective suicide rates for the years 2014 to 2016. We were investigating potential connections between suicide rates and a country's socioeconomic status.

``` {r, echo=FALSE}
chart2
```

However, as we can observe from this scatterplot, there does not appear to be a correlation between the GDP per capita of a country and the suicide rate of the country across all 3 years. The hi


## Conclusion

